---
title: "CPBMI R Ch7. Probability Distribution, Hypothesis Testing"
author: "JW Yi"
date: "Monday, September 29, 2014"
output:
  html_document:
    highlight: pygments
    theme: united
---

# Probability Distribution Function of Normal Distribution : dnorm()
## dnorm(z, mean=0, sd=1)
```{r}
z <- seq(-4, 4, by=0.01)
plot(z, dnorm(z), type="l")
z <- 1.96
dnorm(z)
abline(h=dnorm(z), v=z, col="red")
```

# Cumulative Distribution Function of Normal Distribution : pnorm()
## pnorm(1.96)
```{r}
pnorm(1.96)
z <- seq(-4, 4, by=0.01)
plot(z, dnorm(z), type="l")
z <- c( -4, seq(-4, 1.96, by=0.01), 1.96 )
fz <- c( 0, dnorm( seq(-4, 1.96, by=0.01) ), 0 ) 
polygon(z, fz, col="red", density=c(10))
```
## pnorm(-1.96)
```{r}
pnorm(-1.96)
z <- seq(-4, 4, by=0.01) 
plot(z, dnorm(z), type="l" )
z <- c( -4, seq(-4, -1.96, by=0.01), -1.96 )
fz <- c( 0, dnorm( seq(-4, -1.96, by=0.01) ), 0 ) 
polygon(z, fz, col="red", density=c(10) )
```

## 1- pnorm()
```{r}
1 - pnorm(1.96)
z <- seq(-4, 4, by=0.01)
plot(z, dnorm(z), type="l")
z <- c(1.96, seq(1.96, 4, by=0.01), 4)
fz <- c(0, dnorm(seq(1.96, 4, by=0.01)), 0) 
polygon(z, fz, col="red", density=c(10))
```

# Inverse Normal Function : qnorm()
## qnorm()
```{r}
qnorm(0.025)
qnorm(0.975)
```

# Hypothesis Testing
## p-value
```{r}
2 * pnorm(-1.41)
```

## One-sample t-test
```{r}
intake <- c(5260, 5470, 5640, 6180, 6390, 6515, 6805, 7515, 7515, 8230, 8770)
mu_0 <- 7725
intake.t <- (mean(intake) - mu_0)/(sd(intake)/sqrt(length(intake)))
intake.t
2 * pt(intake.t, df=length(intake)-1)
```

## Plot
```{r}
tx <- seq(-3, 3, by=0.01)
plot(tx, dt(tx, 10), type="l")
qt(0.025, df=10)
abline(v= -2.228139, col="red")
abline(v= 2.228139, col="red")
```

## t.test(samples, mu=mu_0)
```{r}
mu_0 <- 7725
t.test(intake, mu = mu_0)
intake.t.test <- t.test(intake, mu = mu_0)
names(intake.t.test)
intake.t.test$statistic
intake.t.test$p.value
intake.t.test$conf.int
intake.t.test$conf.int[1]
intake.t.test$conf.int[2]
```

# Binomial Distribution  
## dbinom(x, size=..., prob=...)
```{r}
dbinom(14, size=20, prob=0.5)
```

## pbinom(x, size=..., prob=...)
```{r}
1-pbinom(10, size=26, prob=1/4)
x <- 0:26
plot(x, dbinom(x, size=26, prob=1/4), type="h")
1-sum(dbinom(0:10, size=26, prob=1/4))
```

# Chi-Square Distribution
## dchisq()
```{r}
x <- seq(1, 50, 0.05)
plot(x, dchisq(x, 1), type="l")
lines(x, dchisq(x, 5), type="l", col="red") 
lines(x, dchisq(x, 10), type="l", col="blue") 
lines(x, dchisq(x, 20), type="l", col="green") 
lines(x, dchisq(x, 40), type="l", col="cyan")
color_list <- c("black", "red", "blue", "green", "cyan") 
legend("topright", c("df=1", "df=5", "df=10", "df=20", "df=40"), 
     lty=c(1,1,1,1,1), col=color_list)

1-pchisq(7.977, 4) # p-value
```

# F-Distribution
## df(), pf(), qf()
```{r}
qf(0.99, 2, 12)
1-pf(15.5, 2, 12)
```
