---
title: "CPBMI R Ch4. Advanced Data Manipulation"
author: "JW Yi"
date: "Thursday, September 25, 2014"
output:
  html_document:
    highlight: pygments
    theme: united
---

# Merge Data
```{r}
temperature <- read.table("D://Work/CPBMI/temperature.txt", header=TRUE, sep=" ")
humidity <- read.table("D://Work/CPBMI/humidity.txt", header=FALSE, na.strings=c(".", "NA"), sep=",")
colnames(humidity) <- c("area", "humidity")
head(temperature)
head(humidity)
```
## Inner join
```{r}
merge(temperature, humidity, by=c("area"))
```
## Full outer join
```{r}
merge(temperature, humidity, by=c("area"), all=TRUE)
```
## Left outer join 
```{r}
merge(temperature, humidity, by=c("area"), all.x=TRUE)
```
## Right outer join 
```{r}
merge(temperature, humidity, by=c("area"), all.y=TRUE)
```

# Package Installation : reshape2, dplyr packages for advanced data manipulation
```{r}
#install.packages("reshape2")
#install.packages("dplyr")
```

# Reshape2 package
## melt() : Convert "Wide format" data into "Long format"
```{r}
library(reshape2)
dat.wide <- read.table("D://Work/CPBMI/dat_wide.txt", header=TRUE, sep="\t")
dat.wide
melt(dat.wide, id.var = c("age", "sex"))
dat.long <- melt(dat.wide, id.var=c("age","sex"), variable.name="condition", value.name="score")
dat.long
```
## dcast() : Revert "Long format" data into "Wide format"
With aggregation
```{r}
dcast(dat.long, age ~ condition, mean, value.var = "score")
dcast(dat.long, sex ~ condition, mean, value.var = "score")
dcast(dat.long, age ~ sex, mean, value.var = "score")
```
Without aggregation
```{r}
dcast(dat.long, age + sex ~ condition, value.var = "score")
dcast(dat.long, age + condition ~ sex, value.var = "score")
dcast(dat.long, age ~ condition + sex, value.var = "score")
```

# dplyr package
```{r}
library(MASS)
library(dplyr)

birthwt_df <- tbl_df(birthwt) # tbl_df : easy looking dataframe display
birthwt_df
```
## select() : Make an advanced selection
```{r}
select(birthwt_df, low, age, smoke, bwt) # Select specific columns
select(birthwt_df, -c(ptl, ht, ui, ftv)) # Select columns except -c(col1, col2, ...)
select(birthwt_df, low:smoke)   # Column selection with indexing
```
## filter(), arrange(), mutate()
```{r}
filter(birthwt_df, age == 21, smoke == 1)
arrange(birthwt_df, age, bwt)
arrange(birthwt_df, desc(age), desc(bwt))
mutate(birthwt_df, bwt_kg = bwt/1000)
```
## group_by(), summarise()
```{r}
summarise(birthwt_df, bwt_mean = mean(bwt, na.rm=TRUE), bwt_median = median(bwt))
group_smoke <- group_by(birthwt_df, smoke)
summarise(group_smoke, count=n(), bwt_mean = mean(bwt), bwt_median = median(bwt)) # n() : counting numbers
```
## Chain operator : %.%
```{r}
birthwt_df %.%
filter(age >= 21 & age <= 30) %.%
group_by(race, smoke) %.%
summarise(count=n(), bwt_mean=mean(bwt, na.rm=TRUE))
```
