---
title: "CPBMI R Ch8. Basic Statistical Tests"
author: "JW Yi"
date: "Monday, September 29, 2014"
output:
  html_document:
    highlight: pygments
    theme: united
---

# Check Normality : shapiro.test()
```{r}
library(MASS)
attach(Pima.tr)
shapiro.test(bmi) 
 # H0 : This data follow normal distribution
 # H1 : This data not follow normal distribution
```
This data follow normal distribution (p > 0.05)
```{r}
qqnorm(bmi) # Scatter Plot
qqline(bmi) # Normality Line
```

# One-Sample T-Test
```{r}
t.test(bmi, mu=30)
t.test(bmi, mu=30, alternative="less")
t.test(bmi, mu=30, alternative="greater")
t.test(bmi, mu=30, conf.level=0.9)
```

# Variance Test : var.test()
```{r}
var.test(bmi ~ type)
```
Variance is different between two samples. => Welch t-test

# Two-Sample T-Test : t.test(data1, data2) or t.test(data ~ variable)
```{r}
t.test(bmi ~ type)
```
Same code
```{r}
Pima.tr.no <- Pima.tr$bmi[Pima.tr$type=="No"]
Pima.tr.yes <- Pima.tr$bmi[Pima.tr$type=="Yes"]
t.test(Pima.tr.no, Pima.tr.yes)
```

# Paired T-Test : t.test(data1, data2, paired=TRUE)
```{r}
attach(anorexia)
boxplot(Prewt - Postwt ~ Treat, col=rainbow(3))
summary(anorexia)
FT <- anorexia[anorexia$Treat=="FT",]
head(FT)
shapiro.test(FT$Prewt - FT$Postwt)
t.test(FT$Prewt, FT$Postwt, paired=TRUE)
```

# Non-Parametric Test 
## Wilcoxon Signed Rank Test : wilcox.test(data1, data2, paired = TRUE)
```{r}
anorexia.cbt <- anorexia[anorexia$Treat=="CBT", ]
shapiro.test(anorexia.cbt$Prewt - anorexia.cbt$Postwt)
wilcox.test(anorexia.cbt$Prewt , anorexia.cbt$Postwt, paired=TRUE)
```
## Wilcoxon Rank-Sum Test(Mann-Whitney U test) : wilcox.test(data1, data2)
```{r}
exposure <- c(14.4, 14.2, 13.8, 16.5, 14.1, 16.6, 15.9, 15.6, 14.1, 15.3, 15.7, 16.7, 13.7, 15.3, 14.0)
non.exposure <- c(17.4, 16.2, 17.1, 17.5, 15.0, 16.0, 16.9, 15.0, 16.3, 16.8)
wilcox.test(exposure, non.exposure)
```

# Chi-Square Test : chist.test()
## Goodness of Fit
```{r}
chisq.test(c(24,16), p=c(0.7,0.3))
```
## Test for Independence
```{r}
countTable <- matrix(c(10845, 189, 10933, 104), nrow=2, byrow=TRUE)
rownames(countTable) <-c("Placebo", "Aspirin")
colnames(countTable) <-c("No Heart Attack", "Heart Attack")
countTable
chisq.test(countTable)

birthwt <- read.table("D://Work/birthwt2.txt", header=TRUE) 
smoke_low_tb <- table(birthwt$smoke, birthwt$low)
smoke_low_tb
chisq.test(smoke_low_tb)
```

# Fisher's Exact Test
```{r}
TeaTasting <- matrix(c(3, 1, 1, 3),  nrow = 2, 
                     dimnames = list( Guess = c("Milk", "Tea"), Truth = c("Milk", "Tea")))
fisher.test(TeaTasting)
```

# Cocran-Armitage Trend Test : 2 * k table
```{r}
prop.trend.test(c(13, 7, 21), c(42, 14, 28))

prop.trend.test(c(13, 10, 10), c(42, 14, 28))
chisq.test(matrix(c(13, 29, 10, 4, 10, 18), ncol=3))
```

# McNemar's Test : binary variable
```{r}
mcnemar.test(matrix(c(5, 15, 5, 7), ncol=2))
```

# CrssTable()
```{r}
library(gmodels) 
CrossTable(countTable, fisher=TRUE, chisq=TRUE, mcnemar=TRUE, expected=TRUE, sresid=TRUE, format="SPSS")
```

# One-Way ANOVA
```{r}
library(MASS)
attach(anorexia)
Change <-  Postwt - Prewt
boxplot(Change ~ Treat, col=rainbow(3))
shapiro.test(Change)
bartlett.test(Change ~ Treat)
aov.out  <- aov(Change ~ Treat)
summary(aov.out)
```

## Post-Hoc Analysis
```{r}
TukeyHSD(aov.out)
plot(TukeyHSD(aov.out))
```

## Two-Way ANOVA without Interaction
```{r}
program1 <- c(9, 12, 14, 11, 13)
program2 <- c(10, 6, 9, 9, 10)
program3 <- c(12, 14, 11, 13, 11)
program4 <- c(9, 8, 11, 7, 8)
dat <- data.frame(min=c(program1, program2, program3, program4),
                  type=c(rep("program1", 5), rep("program2", 5), rep("program3", 5), rep("program4", 5)))
aov.out <- aov(min ~ type, data=dat)
summary(aov.out)
TukeyHSD(aov.out)
plot(TukeyHSD(aov.out))
```
Other Example
```{r}
teaching_time <- read.table("D://Work/teaching_time.txt", header=TRUE, sep=" ")
teaching_time
aov.out <- aov(days ~ ageGroup + method, data=teaching_time)
summary(aov.out)
```

# Two Way Anova with Interaction
```{r}
summary(ToothGrowth)
ToothGrowth$dose <- factor(ToothGrowth$dose)
summary(ToothGrowth)
aov.out <- aov(len ~ supp*dose, data=ToothGrowth)
summary(aov.out)
interaction.plot(ToothGrowth$dose, ToothGrowth$supp, ToothGrowth$len)
```
