---
title: "CPBMI R Ch2. Data Control"
author: "JW Yi"
date: "Tuesday, September 23, 2014"
output:
  html_document:
    highlight: pygments
    theme: united
---

# Matrix Binding
## Row and Column Binding
```{r}
mat <- matrix( 1:20, nrow=4)
mat
cbind(mat, c(21:24)) # Column binding
rbind(mat, c(21:25)) # Row binding
mat <- rbind(mat, c(21:25)) # Assign into new variable
mat
```
## Binding 2 Matrices
```{r}
matA <- matrix(c(1:8), nrow=4)
matB <- matrix(c(20:27), nrow=4)
matA
matB
cbind(matA, matB)
rbind(matA, matB)
```
## rowSums, colSums
```{r}
data <- matrix(c(1:8), nrow=4)
data
rowSums(data)
apply(data, 1, sum)

colSums(data)
apply(data, 2, sum)

rowMeans(data)
apply(data, 1, mean)

colMeans(data)
apply(data, 2, mean)
```

# Factor : Categorical Variable
```{r}
bloodType <- c("A","B", "AB", "O", "O", "A", "A", "O", "B", "B")
summary(bloodType)
```
## Let's change "character" class into "factor"
```{r}
bloodType <- factor(bloodType)
bloodType
summary(bloodType)
```
## Some data have "numeric" coding for "factor" variable. R can control "factor" variable with their own name.
```{r}
gender <- c(1,1,2,2,1,2,2,2,1,2)
summary(gender)
gender <- factor(gender, levels = c(1,2), labels = c("male", "female"))
 # factor(data, levels = c(level 1, level 2, ... ), labels = c("name 1", "name 2", ...))
gender
summary(gender)
```

# Data Frame : Matrix that has multiple types of variables.
```{r}
intake.pre <- c(5260, 5470, 5640, 6180, 6390, 6515, 6805, 7515, 7515, 8230, 8770)
intake.post <- c(3910, 4220, 3885, 5160, 5645, 4680, 5265, 5975, 6790, 6990, 7335)
intake.race <- c('w', 'w', 'b', 'y', 'y', 'w', 'b', 'w', 'w', 'y', 'b')

intake <- data.frame(intake.pre, intake.post, intake.race)
intake
```
## Indexing Dataframe is same as matrix.
```{r}
intake[, 1]
intake$intake.pre
intake[1, ]
```
## We also name the columns and rows.
```{r}
intake <- data.frame( before = intake.pre, after=intake.post, race=intake.race )
intake
rownames(intake) <- c("w1", "w2", "w3", "w4", "w5", "w6", "w7", "w8", "w9", "w10", "w11")
intake
intake['w11', ]
```
## Summary and Check Structures.
```{r}
summary(intake)
str(intake)
```
## Conditional Syntax of Dataframe
```{r}
intake [intake$before > 7000, ]
intake [intake$before > 7000 & intake$before < 8000, ]
intake [intake$before - intake$after > 1300 & intake$race == "w", ]
```

# List 
## List can harbor all data structures in it.
```{r}
list_ex <- list(1:10, c("A","B","C"), matrix(1:4, nrow=2))
list_ex
summary(list_ex)
```
## Indexing list : [[ ]]
```{r}
list_ex [[1]]
list_ex [[2]]
list_ex [[3]]
list_ex [[1]][5]
```
## Naming List
```{r}
list_ex <- list(vector_ex = 1:10, char_ex = c("A","B","C"), matrix_ex = matrix(1:4, nrow=2), 
                df_ex = data.frame(pre=1:10, post=101:110))
list_ex
names(list_ex)
list_ex $ df_ex[1:2, "pre"]
```
## Adding New Element 
```{r}
list_ex <- append(list_ex, list(gender=factor(c("M","F"))))
list_ex
```

# Missing Value : NA
```{r}
y <- c(1,2,3,NA)
y
is.na(y)
summary(y)

y <- factor(c("A","B","C",NA))
y

ages <- c(48, 78, 56, 88, -999, 13, 26, -999)
ages
ages[ages == -999] <- NA
ages
summary(ages)
sum(is.na(ages))
```
## Missing value NA is differ from character string "NA"
```{r}
is.na(NA)
is.na("NA") 
```
## If data have NA, all statistic result also NA.
```{r}
1 + 2 + NA
1 * 2 * NA
NA * NA
NA / 2
NA > 2
sum(ages)
mean(ages)
```
## na.rm = TRUE option can solve this problem.
```{r}
sum(ages, na.rm = TRUE)
mean(ages, na.rm = TRUE)
```
## na.omit(), complete.cases(). 
```{r}
weight <- c(65.4, 55, 380, 72.2, 51, NA)
height <- c(170, 155, NA, 173, 161, 166)
gender <- c("M", "F", "M", "M", "F", "F")

patients <- data.frame(weight=weight, height=height, gender=gender)
patients
str(patients)

na.omit (patients)
complete.cases(patients)
patients[complete.cases(patients), ]
```
## Filtering Missing Values
```{r}
mat <- matrix(1:50, nrow=5)
mat[1,1] <- NA; mat[1,3] <- NA; mat[2, 2:7] <- NA; mat[4,3] <- NA
mat

dim(mat)
na.omit(mat)

mat[rowSums(is.na(mat))/dim(mat)[2] > 0.2, ]
mat[rowSums(is.na(mat))/dim(mat)[2] > 0.1, ]
```
## Adding "na.rm" Arguments to Apply Function
```{r}
patients.sub <- patients[ ,c("weight", "height")]
patients.sub
apply(patients.sub, 2, mean)
apply(patients.sub, 2, mean, na.rm=TRUE)
```

# Read Data From Files
## Read Text Files
```{r}
# myData <- read.table(file, header=TRUE, sep="\t", row.names=1, na.strings = c("NA", "Not Applicable", ...))
temperature <- read.table("temperature.txt", header=TRUE, sep=" ")
temperature
summary(temperature)

humidity <- read.table("humidity.txt", header=FALSE, sep=",", na.strings=c(".","NA"))
humidity
colnames(humidity) <- c("area", "humidity")
head(humidity, n=2)
summary(humidity)
```
## Write Data into Text Files
```{r}
# write.table(data, file="file name", sep="seperator", quote=FALSE, row.names=TRUE, ...)
tempF <- temperature[ ,"temperature"] * 9/5 + 32
tempF

cbind(temperature, data.frame(tempF = tempF))
temperature <- cbind(temperature, data.frame(tempF = tempF))

write.table(temperature, file="tempF.txt", sep="\t", quote=FALSE, row.names=FALSE)
```
## Write and Read Data into RDS files
```{r}
# saveRDS(data, "filename.rds")
# dat <- readRDS("filename.rds")
```
