---
output:
  html_document:
    highlight: pygments
    theme: united
---
CPBMI R CH1. Introduction of R
==============================

# Basic Introduction
```{r}
2+2
exp(-3)
rnorm(10)

x <- 30
x

weight <- 78.5
Weight <- 110.3
weight
Weight
```

# " <- " vs " = "
```{r}
divide <- function(A,B) {A/B}
divide (1,2)
divide (2,1)

divide (A = 2, B = 1)
divide (A <- 2, B <- 1)
divide (B = 2, A = 1)
divide (B <- 2, A <- 1)
```

# Storage mode
## Numeric
```{r}
value <- 605
value
mode(value)
```
## Character
```{r}
string <- "Hello CPBMI R"
string
mode(string)

string_no <- "605"
string_no
mode(string_no)
```
## Logical
```{r}
logic <- 2<4
logic
mode(logic)
```
## Complex Number
```{r}
compx <- 2 + 3i
compx
mode(compx)
```

# Vector
```{r}
x1 <- c(1,2,3,4)
x1
length(x1)

x2 <- c("A","B","C")
x2
length(x2)

x3 <- c(1,2,"A","B")
x3
length(x3)
is.vector(x3)
```
## Creating Vector
```{r}
rep(2,2)
rep(c(2,3), each=2)
seq(0,1, length=11)
seq(1,9, by=2)
c(1:10)
1:10
```
## Vectorized Arithmetic
```{r}
weight <- c(60, 72, 57, 90, 95, 72)
height <- c(1.75, 1.80, 1.65, 1.90, 1.74, 1.91)
bmi <- weight / height ^ 2
bmi
```
## Simple Statistics
```{r}
min(weight)
min(height)
max(weight)
max(height)
range(weight)
range(height)
mean(weight)
mean(height)
median(weight)
median(height)
var(weight)
var(height)
sd(weight)
sd(height)
cor(weight, height)
quantile(weight)
quantile(height)
summary(weight)
summary(height)
str(weight)
str(height)
```
## Indexing
```{r}
x <- 11:20
x
x[5]
x[1:4]
x[5] <- 99
x[5]
x[c(1,3,5)]
```
## Logical Syntax
```{r}
x > 15
x[x>15]
which(x>15)
x[which(x>15)]
which(x == 99)
```

# Matrix
```{r}
y <- matrix(1:20, nrow=5, ncol=4)
y

y <- matrix(1:20, nrow=5, ncol=4, byrow=TRUE)  # defalt byrow=FALSE
y

dim(y)

t(y) # transpose matrix
```
## Indexing Matrix
```{r}
y[3,2]
y[2,4]
y[,1]
y[1,]
y[c(3,4),]
y[3,2] <- 99
```
## Changing Element
```{r}
y
y[5,] <- c(117,118,119,120)
y
y[,3] <- c(50,51,52,53,54)
y
y[4, c(3,4)] <- c(115,116)
y
```
## Conditional Syntax
```{r}
y[y>100]
which(y>100)
which(y>100, arr.ind=TRUE)
y[which(y>100, arr.ind=TRUE)]
```
## Matrix Calculation
```{r}
x <- matrix(c(1,2,3,4), nrow=2)
y <- matrix(c(5,6,7,8), nrow=2)
x
y
x + y
x - y 
x * y
x %*% y
x / y
```

# Naming
## Naming Vector
```{r}
x <- c(10,30,40,50)
x
names(x) <- c("A","B","C","D")
x
names(x)
x["B"]
x[2]
```
## Naming Matrix Row and Column
```{r}
countTable <- matrix(c(189,10845,104,10933), nrow=2, byrow=TRUE)
countTable
rownames(countTable) <- c("Placebo", "Aspirin")
colnames(countTable) <- c("Heart Attack", "No Heart Attack")
countTable
countTable["Placebo",]
countTable["Aspirin",]
countTable[, "No Heart Attack"]
countTable[, "Heart Attack"]
```

# Apply : apply(matrix, margin - 1(row) or 2(col), function)
```{r}
X <- matrix(1:20, nrow=4)
X
apply(X, 1, sum) 
apply(X, 2, mean)
```

