---
title: "CPBMI R Ch3."
author: "JW Yi"
date: "Wednesday, September 24, 2014"
output:
  html_document:
    highlight: pygments
    theme: united
---

# Attach, Detach, Paths
## ls(), attach()
```{r}
ls()
a <- c(1,2,3)
b <- data.frame(id=c(1,2,3), value=c("A","B","C"))
ls()
b$value

attach(b)
value
```
## search(), searchpaths()
```{r}
search()
searchpaths()
```
## detach(), rm()
```{r}
detach(b)
search()

ls()
rm(a)
ls()
```
## Masking 
```{r}
df1 <- data.frame(id = c("A","B","C"), cost=c(1.2,3,2.2))
df2 <- data.frame(id=c("w1", "w2", "w3"), cost=c(2.2, 5, 1.2))

attach(df1)
attach(df2)
id
cost
```

# with() : with( object , statement )
```{r}
with(df1, summary(id))
with(df1, {
     summary(id);
     summary(cost);
})

with(df1, {
     summary.id <<- summary(id)  # <<- : Global assignment operator
     summary.cost <- summary(cost)
})
summary.id
# summary.cost : Error: object 'summary.cost' not found
```
# Type Check and Conversion
```{r}
df <- data.frame(id = LETTERS[1:10], id2 = letters[1:10], val1 = 1:10, val2 = 1:10)
df
is.data.frame(df)

df.matrix <- as.matrix(df)
df.matrix
```

# Woking With Strings
## Get the Length
```{r}
x <- c("normal_1", "normal_2", "cancer_1", "cancer_2")
length(x) # Get the length of vector
nchar(x[1]) # Get the length of element
```
## Upper / Lower case conversion
```{r}
toupper(x)
tolower(x)
```
## Select Specific Letters
```{r}
substr(x, 1, 6)

grep("normal", x, ignore.case=TRUE)  # ignore.case=FALSE : distinguish upper/lower case letters
grep("cancer", x, ignore.case=TRUE)  # fixed=FALSE : regular expression
```
## Escape letter : \\
```{r}
x <- "Hello CPBMI"
strsplit(x, "\\s")
unlist(strsplit(x, "\\s"))
```
## Substitution
```{r}
sub("CPBMI", "CPBMI2014", x)
```
## Paste
```{r}
chisq_value <- 8.07
degree_f <- 1
pvalue <- 0.0045
paste("chisq =", chisq_value, "df =", degree_f, "p-value =", pvalue, sep=" ")
```

# if / else / ifelse
```{r}
birthwt <- read.table("D://Work/birthwt.txt", header=T, sep=" ")
if ( !is.factor(birthwt$smoke) ) 
{ 
     birthwt$smoke <- factor(birthwt$smoke) 
} else { 
	print("Smoke is factor.")
}

if ( !is.factor(birthwt$smoke) ) 
{ 
     birthwt$smoke <- factor(birthwt$smoke) 
} else { 
     print("Smoke is factor.")
}

ifelse(birthwt$low == 0, "low", "high")
scores <- c(55, 70, 80, 90, 59, 87)

ifelse (scores >= 60, "Pass", "Fail")
```

# For loop
## Basic concept for "for loop"
```{r}
for (i in 1:5) print (i)
for (i in 2:9) {
     for (j in 1:9){
          print (paste (i, "*", j, "=", i*j, sep=""))
     }}
```
## for loop vs apply 
```{r}
mat <- matrix(rnorm(50000 * 100), nrow=50000)
mat.rowSum <- c()
system.time(for (i in 1:50000){mat.rowSum[i] <- sum(mat[i, ])})
system.time(sumMat.apply <- apply(mat, 1, sum))
```
"Apply" is faster than "for loop".

# Function
```{r}
mymean <- function(x) { 
     my.mean <- mean(x)
     return(my.mean)}
a <- c(1,2,3,4,5)
mymean(a)

a <- c(1,2,3,4,5,NA)
mymean(a)

mymean <- function(x, remove.na = FALSE) {
     my.mean <- mean(x, na.rm = remove.na)
     return(my.mean)
}
mymean(a)
mymean(a, remove.na = TRUE)

BMI <- function(weight, height){
     if (weight > 300 | height > 300) {
          stop ("Too high for weight or height")}
     return(weight / (height/100)** 2)
}
BMI (74, 174)

mystats <- function(x, print = FALSE, ...){
     my.mean <- mean(x, ...)
     my.median <- median(x, ...)
     my.sd <- sd(x, ...)
     if (print == TRUE){
          paste ("mean =", my.mean, "median =", my.median, "sd =", my.sd, sep = " ")
     }
     else { res <- list(mean = my.mean, median = my.median, sd = my.sd)
            return (res)}
}
res <- mystats(birthwt$bwt)
res
res <- mystats(birthwt$bwt, print = TRUE, na.rm = TRUE)
res
```

# Apply : lapply(data, col/row, function), sapply(data, col/row, function), tapply(data, group, function)
```{r}
lapply(birthwt[, c("lwt", "bwt")], mean)
sapply(birthwt[, c("lwt", "bwt")], mean)
tapply(birthwt$bwt, birthwt$race, mean)

test <- list(a=1:10, b=c("A","B","C"))
test
unlist(test)
unlist(lapply(birthwt[,c("lwt","bwt")], mean))
sapply(apply(birthwt[,c("lwt","bwt")], 2, mystats), unlist)
```

# Subset : subset(df, conditions..., select=c("col1","col2",...))
```{r}
attach(birthwt)
head(birthwt[(age >25 & bwt > 2500), ])
head(subset(birthwt, age >25 & bwt > 2500))
head(subset(birthwt, age>25 & bwt>2500, select = c("age", "smoke", "bwt")))
```

# Transform : transform(df, column name 1=variable 1, column name 2=variable 2, ...)
```{r}
head(cbind(birthwt, weight_kg = birthwt$bwt/1000))
head(transform(birthwt, weight_kg = bwt/1000))
head(transform(birthwt, weight_kg = bwt/1000, weight_pound = 2.2*bwt))
```

# Order, sort
```{r}
x <- c(2,5,4,8,6)
sort(x) # arrange values
order(x) # return index of each value

ch2_birthwt <- read.table("birthwt.txt", header=TRUE, sep=" ")
head(ch2_birthwt[order(ch2_birthwt$bwt),])
head(ch2_birthwt[order(-ch2_birthwt$bwt),])
head(ch2_birthwt[order(-ch2_birthwt$race, -ch2_birthwt$bwt),])
```

# Grouping Calculation : aggregate(target column, by=list(col1, col2, ...), function)
```{r}
aggregate(bwt, by=list(race, smoke), mean, data=ch2_birthwt)
aggregate(list(bwt.mean=bwt), by=list(race=race, smoke=smoke), mean, data=ch2_birthwt)
```
